<!-- system_diversity.Rmd -->

---
title: "Descriptive overview of LLM extractions: run `r run`"
author: "Zhongtian Sun, Maxwell J. Farrell, Liam Brierley"
date: "`r format(Sys.time(), '%B %d %Y')`"
output: 
html_document:
highlight: default
toc: yes
toc_depth: 3
theme: yeti
urlcolor: blue
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, error=TRUE)
options(kableExtra.html.bsTable = T)

custom_kable <- function(x, caption=NULL) {
  kable(x, caption=caption) %>% kable_styling(font_size = 10, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = T)
}
```

```{r loading packages, echo=F, message=F, warning=F}
## Packages
require(tidyverse)
require(kableExtra)
require(ggplot2)
require(magrittr)
```

```{r reading data, echo=F, message=F, warning=F}
setwd(paste0("H:\\Working\\hp-extractor\\zs\\", run))

files <- list.files()

ent_d <- read.csv(files[grepl("entities", files) & grepl("deepseek", files)]) %>% distinct
ent_g <- read.csv(files[grepl("entities", files) & grepl("gemma", files)]) %>% distinct
ent_l <- read.csv(files[grepl("entities", files) & grepl("llama3-8b|llama.csv", files)]) %>% distinct
ent_ll <- read.csv(files[grepl("entities", files) & grepl("llama3-70b", files)]) %>% distinct
ent_m <- read.csv(files[grepl("entities", files) & grepl("mistral", files)]) %>% distinct

rel_d <- read.csv(files[grepl("relationships", files) & grepl("deepseek", files)])%>% distinct
rel_g <- read.csv(files[grepl("relationships", files) & grepl("gemma", files)]) %>% distinct
rel_l <- read.csv(files[grepl("relationships", files) & grepl("llama3-8b|llama.csv", files)]) %>% distinct
rel_ll <- read.csv(files[grepl("relationships", files) & grepl("llama3-70b", files)]) %>% distinct
rel_m <- read.csv(files[grepl("relationships", files) & grepl("mistral", files)]) %>% distinct

meta_aug <- read.table("H:\\Working\\system_specificity\\raw_data\\dat_aug9_2021.tsv", sep = '\t', comment.char="", quote = "\"", header=TRUE) %>% distinct

val <- read.csv("H:\\Working\\hp-extractor\\raw_data\\validation_100.csv")

```

#### Check all 100 docs are present in each file:

```{r 100check, echo=F, message=F, warning=F}

ent_d %>% pull(doc_id) %>% unique %>% length
ent_g %>% pull(doc_id) %>% unique %>% length
ent_l %>% pull(doc_id) %>% unique %>% length
ent_m %>% pull(doc_id) %>% unique %>% length
rel_d %>% pull(doc_id) %>% unique %>% length
rel_g %>% pull(doc_id) %>% unique %>% length
rel_l %>% pull(doc_id) %>% unique %>% length
rel_m %>% pull(doc_id) %>% unique %>% length


```
&nbsp;
&nbsp;

#### Are there duplicated rows in any model outputs?

_Should be zero._

```{r duprows, echo=F, message=FALSE, layout="l-body-outset"}
data.frame("Duplicated entitity rows" = 
             c(ent_d %>% group_by(entity, type, doc_id) %>% tally %>% filter(n > 1) %>% nrow,
               ent_g %>% group_by(entity, type, doc_id) %>% tally %>% filter(n > 1) %>% nrow,
              ent_l %>% group_by(entity, type, doc_id) %>% tally %>% filter(n > 1) %>% nrow,
               ent_ll %>% group_by(entity, type, doc_id) %>% tally %>% filter(n > 1) %>% nrow,
              ent_m %>% group_by(entity, type, doc_id) %>% tally %>% filter(n > 1) %>% nrow),
           "Duplicated relation rows" = 
             c(rel_d %>% group_by(source_text, target_text, relation_type, doc_id) %>% tally %>% filter(n > 1) %>% nrow,
               rel_g %>% group_by(source_text, target_text, relation_type, doc_id) %>% tally %>% filter(n > 1) %>% nrow,
              rel_l %>% group_by(source_text, target_text, relation_type, doc_id) %>% tally %>% filter(n > 1) %>% nrow,
               rel_ll %>% group_by(source_text, target_text, relation_type, doc_id) %>% tally %>% filter(n > 1) %>% nrow,
              rel_m %>% group_by(source_text, target_text, relation_type, doc_id) %>% tally %>% filter(n > 1) %>% nrow),
           check.names=FALSE) %>%
  set_rownames(c("deepseek", "gemma2", "llama","llama70b","mistral")) %>%
  knitr::kable()
```
&nbsp;
&nbsp;

### Average entities/relations per doc

__


```{r entrelperdoc, echo=F, message=FALSE,  fig.height=3.0, fig.width=9.0, layout="l-body-outset", results="asis"}

# Avg ents/rels per doc
bind_rows(
  ent_d %>% group_by(doc_id) %>% tally %>% mutate(model = "deepseek"),
  ent_g %>% group_by(doc_id) %>% tally %>% mutate(model = "gemma2"),
  ent_l %>% group_by(doc_id) %>% tally %>% mutate(model = "llama"),
  ent_ll %>% group_by(doc_id) %>% tally %>% mutate(model = "llama-70b"),
  ent_m %>% group_by(doc_id) %>% tally %>% mutate(model = "mistral"),
) %>%
  ggplot(aes(x=n, fill=model)) +
  geom_density(alpha=0.2, position = 'identity') +
  xlab("entities per abstract") +
  theme_bw()


bind_rows(
  rel_d %>% group_by(doc_id) %>% tally %>% mutate(model = "deepseek"),
  rel_g %>% group_by(doc_id) %>% tally %>% mutate(model = "gemma2"),
  rel_l %>% group_by(doc_id) %>% tally %>% mutate(model = "llama"),
  rel_ll %>% group_by(doc_id) %>% tally %>% mutate(model = "llama-70b"),
  rel_m %>% group_by(doc_id) %>% tally %>% mutate(model = "mistral"),
) %>%
  ggplot(aes(x=n, fill=model)) +
  geom_density(alpha=0.2, position = 'identity') +
  xlab("relations per abstract") +
  theme_bw()

```

&nbsp;
&nbsp;

### Most common types of entities

**Note only plotted if a given threshold PER MODEL**

__


```{r typesent, echo=F, message=FALSE,  fig.height=5.0, fig.width=9.0, layout="l-body-outset", results="asis"}

# Types of ent/rel
plot_lev_ent <- bind_rows(
  ent_d %>% group_by(type) %>% tally %>% mutate(model = "deepseek"),
  ent_g %>% group_by(type) %>% tally %>% mutate(model = "gemma2"),
  ent_l %>% group_by(type) %>% tally %>% mutate(model = "llama"),
  ent_ll %>% group_by(type) %>% tally %>% mutate(model = "llama-70b"),
  ent_m %>% group_by(type) %>% tally %>% mutate(model = "mistral"),
) %>% filter(n>25) %>% select(-model) %>% group_by(type) %>% summarise(ov = sum(n)) %>% arrange(ov) %>% pull(type)

bind_rows(
  ent_d %>% group_by(type) %>% tally %>% mutate(model = "deepseek"),
  ent_g %>% group_by(type) %>% tally %>% mutate(model = "gemma2"),
  ent_l %>% group_by(type) %>% tally %>% mutate(model = "llama"),
  ent_ll %>% group_by(type) %>% tally %>% mutate(model = "llama-70b"),
  ent_m %>% group_by(type) %>% tally %>% mutate(model = "mistral"),
) %>%
  filter(n > 50) %>%
  ggplot(aes(y=factor(type, levels = plot_lev_ent), x=n,color=model)) +
  geom_point(size = 5) +
  ylab("entity type (min = 50)") +
  theme_bw()
```

```{r diagnosetypesent, echo=T, message=FALSE}

ent_m %>% filter(type == "Unknown") %>% slice(1:25)

ent_d %>% filter(type == "Undefined") %>% slice(1:25)
```
&nbsp;
&nbsp;

### Compare entity types for same entity texts WITHIN models

_How many entity types is each entity text associated with within each model? 6-7% of entities have >1 type. Ideal is zero?_

```{r entitywithinmodel, echo=F, message=FALSE, layout="l-body-outset"}
bind_rows(ent_d %>% group_by(entity) %>% summarise(nt = n_distinct(type)) %>% with(., table(nt)) %>% prop.table() %>% round(3)*100,
          ent_g %>% group_by(entity) %>% summarise(nt = n_distinct(type)) %>% with(., table(nt)) %>% prop.table() %>% round(3)*100,
          ent_l %>% group_by(entity) %>% summarise(nt = n_distinct(type)) %>% with(., table(nt)) %>% prop.table() %>% round(3)*100,
          ent_ll %>% group_by(entity) %>% summarise(nt = n_distinct(type)) %>% with(., table(nt)) %>% prop.table() %>% round(3)*100,
          ent_m %>% group_by(entity) %>% summarise(nt = n_distinct(type)) %>% with(., table(nt)) %>% prop.table() %>% round(3)*100
) %>% bind_cols(model = c("deepseek", "gemma2", "llama", "llama70b", "mistral"),.) %>%
  knitr::kable()
```

_Example entities with >1 type within model:_

```{r entitywithinmodel_eg, echo=T, message=FALSE, layout="l-body-outset"}
ent_l %>% filter(entity %in% (ent_l %>% group_by(entity) %>% summarise(nt = n_distinct(type)) %>% filter(nt>1) %>% pull(entity))) %>% arrange(desc(entity)) %>% select(-entity_id, -doc_id) %>% distinct %>% slice(1:20)
```

&nbsp;
&nbsp;

### Compare entity types for same entity texts BETWEEN models

_ONLY for entities having a single type within models (i.e., 1's in table above). Rows are n models the entity was present in, columns are n types the entity had between models. The ideal is x = 5, y = 1, i.e., the entity was extracted by all five models and all models agreed on the entity type_

```{r entitybetweenmodel, echo=F, message=FALSE, layout="l-body-outset"}
singletype_entity_ext<- function(df){
  df %>% 
    filter(entity %in% 
             (df %>% group_by(entity) %>% summarise(nt = n_distinct(type)) %>% filter(nt == 1) %>% pull(entity)
             )) %>%
    select(entity, type) %>%
    distinct()
}

base_ent <- bind_rows(
  singletype_entity_ext(ent_d),
  singletype_entity_ext(ent_g),
  singletype_entity_ext(ent_l),
  singletype_entity_ext(ent_ll),
  singletype_entity_ext(ent_m)
) %>% 
  select(entity) %>% 
  distinct

ent_type_comp <- purrr::reduce(
  list(base_ent,
       singletype_entity_ext(ent_d),
       singletype_entity_ext(ent_g),
       singletype_entity_ext(ent_l),
       singletype_entity_ext(ent_ll),
       singletype_entity_ext(ent_m)
  ),
  left_join, by = 'entity') %>%
  magrittr::set_colnames(c("entity","deepseek","gemma2","llama","llama70b","mistral")) %>% 
  mutate(models_present = 5-rowSums(is.na(.))) %>%
  rowwise %>% mutate(n_entity_types = length(unique(na.omit(c(deepseek, gemma2, llama, llama70b, mistral))))) %>%
  ungroup()

ent_type_comp %>%
  write.csv(paste0(run, "//model_entity_type_comparison_", run, ".csv"))

# ent_type_comp %>% with(., table(models_present)) %>% rev
ent_type_comp %>% with(., table(models_present, n_entity_types))%>% .[5:1,] %>% knitr::kable()
```

_Example entity with varying and non-varying types between models. Full entity-by-entity sheet is in file ***model_entity_type_comparison.csv***_

```{r entitybetweenmodel_eg, echo=T, message=FALSE}
ent_type_comp %>% filter(models_present == 5 & n_entity_types == 1) %>% sample_n(1)
ent_type_comp %>% filter(models_present == 5 & n_entity_types == 5) %>% sample_n(1)
```

&nbsp;
&nbsp;

### Most common types of relations

__

```{r typesrel, echo=F, message=FALSE,  fig.height=5.0, fig.width=9.0, layout="l-body-outset", results="asis"}
plot_lev_rel <- bind_rows(
  rel_d %>% group_by(relation_type) %>% tally %>% mutate(model = "deepseek"),
  rel_g %>% group_by(relation_type) %>% tally %>% mutate(model = "gemma2"),
  rel_l %>% group_by(relation_type) %>% tally %>% mutate(model = "llama"),
  rel_ll %>% group_by(relation_type) %>% tally %>% mutate(model = "llama-70b"),
  rel_m %>% group_by(relation_type) %>% tally %>% mutate(model = "mistral"),
) %>% filter(n>25) %>% select(-model) %>% group_by(relation_type) %>% summarise(ov = sum(n)) %>% arrange(ov) %>% pull(relation_type)

bind_rows(
  rel_d %>% group_by(relation_type) %>% tally %>% mutate(model = "deepseek"),
  rel_g %>% group_by(relation_type) %>% tally %>% mutate(model = "gemma2"),
  rel_l %>% group_by(relation_type) %>% tally %>% mutate(model = "llama"),
  rel_ll %>% group_by(relation_type) %>% tally %>% mutate(model = "llama-70b"),
  rel_m %>% group_by(relation_type) %>% tally %>% mutate(model = "mistral"),
) %>%
  filter(n > 25) %>%
  ggplot(aes(y=factor(relation_type, levels = plot_lev_rel), x=n,color=model)) +
  geom_point(size = 5) +
  ylab("relation type (min = 20)") +
  theme_bw()
```
&nbsp;
&nbsp;

### Examine Latin name of/common name of relations

_ALL MODELS have directional errors in common/latin name of, e.g.:. This is probably then easiest to harmonise with RAG or external referencing or even embedding idea_

```{r latincommon, echo=T, message=FALSE, layout="l-body-outset"}
rel_ll %>% filter(relation_type == "common name of") %>% slice(1:10)
```

&nbsp;
&nbsp;

### Abstracts missing predefined entity and relation types:

__

_Seems a mix of possible synonyms for "host" and "infect", e.g., found in" "is affected by", "are hosts of", etc. within the above but need to validate fully to confirm this - IF we do indeed want to do full relation validation (instead of just a priori harmonising)._


```{r missing, echo=F, message=FALSE, layout="l-body-outset"}
data.frame("Abs with no 'Host' entity" = 
             c(ent_d %>% filter(!(doc_id %in% c(ent_d %>% filter(type == "Host") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               ent_g %>% filter(!(doc_id %in% c(ent_g %>% filter(type == "Host") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               ent_l %>% filter(!(doc_id %in% c(ent_l %>% filter(type == "Host") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               ent_ll %>% filter(!(doc_id %in% c(ent_ll %>% filter(type == "Host") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               ent_m %>% filter(!(doc_id %in% c(ent_m %>% filter(type == "Host") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length),
           "Abs with no 'Pathogen' entity" = 
             c(ent_d %>% filter(!(doc_id %in% c(ent_d %>% filter(type == "Pathogen") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               ent_g %>% filter(!(doc_id %in% c(ent_g %>% filter(type == "Pathogen") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               ent_l %>% filter(!(doc_id %in% c(ent_l %>% filter(type == "Pathogen") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               ent_ll %>% filter(!(doc_id %in% c(ent_ll %>% filter(type == "Pathogen") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               ent_m %>% filter(!(doc_id %in% c(ent_m %>% filter(type == "Pathogen") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length),
           "Abs with no 'hosts' relation" = 
             c(rel_d %>% filter(!(doc_id %in% c(rel_d %>% filter(relation_type == "hosts") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               rel_g %>% filter(!(doc_id %in% c(rel_g %>% filter(relation_type == "hosts") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               rel_l %>% filter(!(doc_id %in% c(rel_l %>% filter(relation_type == "hosts") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               rel_ll %>% filter(!(doc_id %in% c(rel_ll %>% filter(relation_type == "hosts") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               rel_m %>% filter(!(doc_id %in% c(rel_m %>% filter(relation_type == "hosts") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length),
           "Abs with no 'infects' relation" = 
             c(rel_d %>% filter(!(doc_id %in% c(rel_d %>% filter(relation_type == "infects") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               rel_g %>% filter(!(doc_id %in% c(rel_g %>% filter(relation_type == "infects") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               rel_l %>% filter(!(doc_id %in% c(rel_l %>% filter(relation_type == "infects") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               rel_ll %>% filter(!(doc_id %in% c(rel_ll %>% filter(relation_type == "infects") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               rel_m %>% filter(!(doc_id %in% c(rel_m %>% filter(relation_type == "infects") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length),
           "Abs with neither relation" = 
             c(rel_d %>% filter(!(doc_id %in% c(rel_d %>% filter(relation_type == "hosts"|relation_type == "infects") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               rel_g %>% filter(!(doc_id %in% c(rel_g %>% filter(relation_type == "hosts"|relation_type == "infects") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               rel_l %>% filter(!(doc_id %in% c(rel_l %>% filter(relation_type == "hosts"|relation_type == "infects") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               rel_ll %>% filter(!(doc_id %in% c(rel_ll %>% filter(relation_type == "hosts"|relation_type == "infects") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length,
               rel_m %>% filter(!(doc_id %in% c(rel_m %>% filter(relation_type == "hosts"|relation_type == "infects") %>% pull(doc_id)))) %>% pull(doc_id) %>% unique %>% length),
           check.names=FALSE) %>%
  set_rownames(c("deepseek", "gemma2", "llama","llama70b","mistral")) %>%
  knitr::kable()
```

__

```{r diagnosemissing, include = FALSE}
rel_ll %>% filter(doc_id %in% (rel_ll %>% filter(!(doc_id %in% c(rel_ll %>% filter(relation_type == "hosts"|relation_type == "infects") %>% pull(doc_id)))) %>% pull(doc_id)))
```

__

```{r diagnosemissing2, include = FALSE}
rel_d %>% filter(relation_type == "no_relationship")
```
&nbsp;
&nbsp;

### Find any relations whose source or target entity do not appear in the matching entities file

_Should be zero for all_

```{r nosourcetarget_process, echo=F}
missing_source_d <- lapply(val$absID, function(i) rel_d %>% filter(doc_id == i & !(tolower(source_text) %in% (ent_d %>% filter(doc_id == i) %>% mutate(entity = tolower(entity)) %>% pull(entity))))) %>%  bind_rows
missing_source_g <- lapply(val$absID, function(i) rel_g %>% filter(doc_id == i & !(tolower(source_text) %in% (ent_g %>% filter(doc_id == i) %>% mutate(entity = tolower(entity)) %>% pull(entity))))) %>%  bind_rows
missing_source_l <- lapply(val$absID, function(i) rel_l %>% filter(doc_id == i & !(tolower(source_text) %in% (ent_l %>% filter(doc_id == i) %>% mutate(entity = tolower(entity)) %>% pull(entity))))) %>%  bind_rows
missing_source_ll <- lapply(val$absID, function(i) rel_ll %>% filter(doc_id == i & !(tolower(source_text) %in% (ent_ll %>% filter(doc_id == i) %>% mutate(entity = tolower(entity)) %>% pull(entity))))) %>%  bind_rows
missing_source_m <- lapply(val$absID, function(i) rel_m %>% filter(doc_id == i & !(tolower(source_text) %in% (ent_m %>% filter(doc_id == i) %>% mutate(entity = tolower(entity)) %>% pull(entity))))) %>%  bind_rows

missing_target_d <- lapply(val$absID, function(i) rel_d %>% filter(doc_id == i & !(tolower(target_text) %in% (ent_d %>% filter(doc_id == i) %>% mutate(entity = tolower(entity)) %>% pull(entity))))) %>%  bind_rows
missing_target_g <- lapply(val$absID, function(i) rel_g %>% filter(doc_id == i & !(tolower(target_text) %in% (ent_g %>% filter(doc_id == i) %>% mutate(entity = tolower(entity)) %>% pull(entity))))) %>%  bind_rows
missing_target_l <- lapply(val$absID, function(i) rel_l %>% filter(doc_id == i & !(tolower(target_text) %in% (ent_l %>% filter(doc_id == i) %>% mutate(entity = tolower(entity)) %>% pull(entity))))) %>%  bind_rows
missing_target_ll <- lapply(val$absID, function(i) rel_ll %>% filter(doc_id == i & !(tolower(target_text) %in% (ent_ll %>% filter(doc_id == i) %>% mutate(entity = tolower(entity)) %>% pull(entity))))) %>%  bind_rows
missing_target_m <- lapply(val$absID, function(i) rel_m %>% filter(doc_id == i & !(tolower(target_text) %in% (ent_m %>% filter(doc_id == i) %>% mutate(entity = tolower(entity)) %>% pull(entity))))) %>%  bind_rows
```

```{r nosourcetarget, echo=F, message=FALSE, layout="l-body-outset"}
data.frame("Abs with 'Source' not in entities" = 
             c(missing_source_d %>% pull(doc_id) %>% unique %>% length,
               missing_source_g %>% pull(doc_id) %>% unique %>% length,
               missing_source_l %>% pull(doc_id) %>% unique %>% length,
               missing_source_ll %>% pull(doc_id) %>% unique %>% length,
               missing_source_m %>% pull(doc_id) %>% unique %>% length),
           "Abs with 'Target' not in entities" = 
             c(missing_target_d %>% pull(doc_id) %>% unique %>% length,
               missing_target_g %>% pull(doc_id) %>% unique %>% length,
               missing_target_l %>% pull(doc_id) %>% unique %>% length,
               missing_target_ll %>% pull(doc_id) %>% unique %>% length,
               missing_target_m %>% pull(doc_id) %>% unique %>% length),
           check.names=FALSE) %>%
  set_rownames(c("deepseek", "gemma2", "llama","llama70b","mistral")) %>%
  knitr::kable()
```

