<!-- manual_validation_display.Rmd -->

---
title: "Manual validation of prompt-1 extracted entities/relations"
author: "Zhongtian Sun, Liam Brierley, Maxwell J. Farrell"
date: "`r format(Sys.time(), '%B %d %Y')`"
output: 
html_document:
highlight: default
toc: yes
toc_depth: 3
theme: yeti
urlcolor: blue
---
<style type="text/css">
.main-container {
  max-width: 100% !important;
  margin: auto;
}
</style>

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, error=TRUE)
options(kableExtra.html.bsTable = T)

custom_kable <- function(x, caption=NULL) {
  kable(x, caption=caption) %>% kable_styling(font_size = 10, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = T)
}
```

## Table to assist reading: highlight entities

```{r table_ent, results="asis"}
ent1 %>%
  left_join(val %>% mutate(absID = as.character(absID)), by = c("doc_id" = "absID")) %>%
  select(-single_multi, -latin_name) %>%
  arrange(factor(doc_id, levels = val$absID)) %>%  # preserve same order as in validation 100 csv
  rowwise() %>%
  mutate(title_abs = gsub(entity, paste0("<b><u>",entity,"</u></b>"), title_abs, ignore.case = TRUE)) %>%
  DT::datatable(escape = FALSE,
            rownames= FALSE,
            options = list(
              autoWidth = TRUE,
              scrollX = TRUE,
              columnDefs = list(list(width = '1500px', targets = 3)))) %>%
  DT::formatStyle(columns = c(1:4), fontSize = '90%')
```

## Table to assist reading: highlight relations

```{r table_rel, results="asis"}
rel1 %>%
  left_join(val %>% mutate(absID = as.character(absID)), by = c("doc_id" = "absID")) %>%
  select(-source_id, -target_id) %>%
  arrange(factor(doc_id, levels = val$absID)) %>%  # preserve same order as in validation 100 csv
  rowwise() %>%
  mutate(title_abs = gsub(source_text, paste0("<b><u>",source_text,"</u></b>"), title_abs, ignore.case = TRUE)) %>%
  mutate(title_abs = gsub(target_text, paste0("<b><u>",target_text,"</u></b>"), title_abs, ignore.case = TRUE)) %>%
  DT::datatable(escape = FALSE,
            rownames= FALSE,
            options = list(
              autoWidth = TRUE,
              scrollX = TRUE,
              columnDefs = list(list(width = '1200px', targets = 5)))) %>%
  DT::formatStyle(columns = c(1:6), fontSize = '90%')
```
